---
title: "Analyzing the Efficiency of Municipal Waste Management"
output:
  html_document:
    df_print: paged
---

Municipal waste management is crucial for sustainable urban development. Efficient waste management can lead to economic savings, reduced environmental impact, and improved public health. With its diverse regions and varied urbanization levels, the target country provides an ideal case study to explore the factors influencing the efficiency of waste management.

table for columns description -

Variable (Label)
region (Region)
province Province
name (Name of municipality)
tc (Cost per capita eur)
cres residual cost per capita
csor Sorted cost per caputa
istat National code
area km2
pop population
alt altitude m.s.l.
isle dummy municipality on isle
sea dummy coastal municipality
pden population density (people per km2)
wden waste per km2
urb urbanization index (1 low, 3 high)
fee fee scheme
d_fee dmmy PAYT
sample Reg with PAYT
organic organic %
paper paper%
glass glass %
wood wood %
metal metal %
plastic plastic %
raee raee %
texile texile %
other other %
msw_so msw sorted kg
msw_un msw unsorted kg
msw Municipal solid waste kg
sor Share of sorther waste
geo 1 South, 2 Center, 3 North
roads Km of roads within the municipality
s_wteregio Share of sw sent to W2E plants - regional figure
s_landfill share of waste to landifil
gdp Municipal revenues EUR (p) - log
proads People per km of roads (log)
wage Taxable income EUR (p) - log
finance Municipal revenues EUR (p) - log


```{r}

library(tidyverse)

setwd("/Users/minervasingh/Documents/Waste")

df=read.csv("public_data_waste_fee.csv")

head(df)

#Understand the structure of the data

str(df)
```

```{r}
summary(df)
```

```{r}
# identify data with missing values
names(which(sapply(df, function(x) any(is.na(x)))))


```
```{r}

df=na.omit(df)

s=summary(df)

```

Correlation between variables

```{r}

# correlation
data_num2 <- select_if(df, is.numeric)             # Subset numeric columns 

# compute correlataion between numeric variables
c= cor(df[, unlist(lapply(df, is.numeric))])    # Properly apply cor  


library(corrplot)

# Insignificant correlation are crossed
col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(x = c, col = col, symm = TRUE)


```


Aims:

(1) To understand the relationship between waste composition, urbanization, population density, and waste management costs in Italian municipalities.
(2) To identify best practices in waste management across different regions and propose recommendations for less efficient municipalities.
 (3) To develop a predictive model that estimates waste management costs based on socio-economic and geographical factors.


Understand the relationship between waste composition, urbanization, population density, and waste management costs in municipalities.

```{r}
  
df$urb=as.factor(df$urb)

df %>% group_by(urb)

```

For urbanisation levels 1, 2 and 3

```{r}

urb1=subset(df,urb=="1")
urb2=subset(df,urb=="2")
urb3=subset(df,urb=="3")

urb1_x=urb1 %>% group_by(region,province,name) %>% 
  summarise(glassm = mean(glass), woodm=mean(wood),metalm=mean(metal),
            platicm=mean(plASTIc),cost=mean(tc),areas=mean(area),gdpm=mean(gdp))

urb1_x

urb2_x=urb2 %>% group_by(region,province,name) %>% 
  summarise(glassm = mean(glass), woodm=mean(wood),metalm=mean(metal),
            platicm=mean(plASTIc),cost=mean(tc),areas=mean(area),gdpm=mean(gdp))

urb2_x

urb3_x=urb3 %>% group_by(region,province,name) %>% 
  summarise(glassm = mean(glass), woodm=mean(wood),metalm=mean(metal),
            platicm=mean(plASTIc),cost=mean(tc),areas=mean(area),gdpm=mean(gdp))

urb3_x

```
```{r}

u1=ggplot(urb1_x, aes(x=glassm, y=gdpm,xName='Cost of Disposing Glass Disturbance',yName='GDP',color=region,legendPosition="bottom")) +geom_point()+ ggtitle("GDP vs Cost of Glass Disposal Across Regions") +
  xlab("Cost of Glass Disposal") + ylab("GDP")

u2=ggplot(urb2_x, aes(x=glassm, y=gdpm,xName='Cost of Disposing Glass Disturbance',yName='GDP',color=region,legendPosition="bottom")) +geom_point()+ ggtitle("GDP vs Cost of Glass Disposal Across Regions") +
  xlab("Cost of Glass Disposal") + ylab("GDP")

u3=ggplot(urb3_x, aes(x=glassm, y=gdpm,xName='Cost of Disposing Glass Disturbance',yName='GDP',color=region,legendPosition="bottom")) +geom_point()+ ggtitle("GDP vs Cost of Glass Disposal Across Regions") +
  xlab("Cost of Glass Disposal") + ylab("GDP")

library(gridExtra)
grid.arrange(u1, u2, u3, nrow = 2)

u1

```

```{r}

u2

```


```{r}

# Selecting numerical columns
numeric_data <-df %>% select_if(is.numeric)

# Scaling the data
scaled_data <- scale(numeric_data)

set.seed(123)
wss <- map_dbl(1:10, function(k){
  kmeans(scaled_data, centers = k, nstart = 10)$tot.withinss
})

# Plot the Elbow Curve
tibble(k = 1:10, wss = wss) %>%
  ggplot(aes(k, wss)) +
  geom_line() +
  geom_point() +
  labs(title = "Elbow Method", x = "Number of clusters", y = "Total within-cluster sum of squares")

library(cluster)    # clustering algorithms # https://uc-r.github.io/kmeans_clustering
library(factoextra)

df2 <- df[,!names(df) %in% c("name", "province","fee")]

x1=df2 %>% group_by(region)%>%summarise_all(list(mean))

library(magrittr)
#https://stackoverflow.com/questions/46165105/error-in-adding-the-first-colum#n-as-the-row-name-in-r
x2=x1 %>% data.frame %>% set_rownames(.$region) %>% select(-region)

k2 <- kmeans(x2, centers = 4, nstart = 25)
k2

fviz_cluster(k2, data = x2)

```


```{r}

x2%>%
  as_tibble() %>%
  mutate(cluster = k2$cluster,
         state = row.names(x2)) %>%
  ggplot(aes(tc, cres, color = factor(cluster), label = state)) +
  geom_text()


```

How do Waste Management Costs Vary Across Municiplaity Clusters?

```{r}

sumtc= x2 %>% mutate(Cluster = k2$cluster) %>% group_by(Cluster) %>%
summarise_all("mean")

p<-ggplot(data=sumtc, aes(x=Cluster, y=tc)) +geom_bar(stat="identity")

p

#https://stackoverflow.com/questions/32468497/creating-barplot-with-standard-errors-plotted-in-r
#https://semba-blog.netlify.app/03/25/2020/barplot-with-errorbar-in-ggplot2/
#https://stackoverflow.com/questions/32468497/creating-barplot-with-standard-errors-plotted-in-r

```

Add standard error bar

```{r}

addcls=sumtc= x2 %>% mutate(Cluster = k2$cluster)

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df3 <- data_summary(addcls, varname="tc",                groupnames=c("Cluster"))

head(df3)

p <- ggplot(df3, aes(x=Cluster, y=tc)) + geom_bar(stat="identity", position=position_dodge(),fill="lightgrey") + geom_errorbar(aes(ymin=tc-sd, ymax=tc+sd), width=.2,position=position_dodge(.9))

p 


```


```{r}

p <- ggplot(df3, aes(x=Cluster, y=tc)) + geom_bar(stat="identity", position=position_dodge(),fill="lightblue") + geom_errorbar(aes(ymin=tc-sd, ymax=tc+sd), width=.2,position=position_dodge(.9))+
    labs(y = "Total Waste Processing Cost")

p 


```




```{r}

dfx= df %>% select(region,urb,fee)

dfx$urb=as.factor(dfx$urb)

head(dfx)

#dx2=dfx %>% data.frame %>% set_rownames(.$name) %>% select(-name)

#head(dx2)
#https://stackoverflow.com/questions/72234213/retain-all-columns-after-using-group-by-summarise-and-mutatue-dplyr-on-categori


```

```{r}

dfx %>% na.omit() %>% group_by(region, urb, fee) %>% dplyr::summarise(ct=n(), .groups = 'drop_last')%>%mutate(proportion = 100 * ct/sum(ct)) %>% filter(region=="ABR") %>% ggplot(aes(x=fee, y=proportion, fill = urb)) + geom_col(position = "dodge")+facet_grid(rows = 'urb') 

```
The fee structure across areas of different levels of urbanisation in a region (ARB)

```{r}

dfx %>% na.omit() %>% group_by(region, urb, fee) %>% dplyr::summarise(ct=n(), .groups = 'drop_last')%>%mutate(proportion = 100 * ct/sum(ct)) %>% filter(region=="ABR") %>% ggplot(aes(x=urb, y=proportion, fill = fee)) + geom_col(position = "dodge")+facet_grid(rows = 'fee') 

```

```{r}

dfx %>% na.omit() %>% group_by(region, urb, fee) %>% 
dplyr::summarise(ct=n(), .groups = 'drop_last') %>% 
mutate(proportion = 100 * ct/sum(ct)) %>% filter(region=="LOM") %>% 
ggplot(aes(x=urb, y=proportion, fill = fee)) + 
geom_col(position = "dodge") +
labs(title = "Distribution of Fee Types Across Urbanization Zones in LOM",
x = "Urbanization Zone", y = "Proportion (%)",fill = "Fee Type")


```


Regression Model For Modelling the Variation in Fees

```{r}

model <- lm(tc ~ pop + area + gdp + paper + texile + glass + wood + metal, data = df)
summary(model)


```

Random forest regression

```{r}

library(randomForest)
library(caret)

set.seed(123)  # Setting a seed ensures reproducibility
splitIndex <- createDataPartition(urb1$tc, p = 0.75, list = FALSE)
train_data <- urb1[splitIndex,]
test_data <- urb1[-splitIndex,]

model1 <- randomForest(tc ~ ., data = train_data, ntree = 100, mtry = 3)
print(model1)

set.seed(123)  # Setting a seed ensures reproducibility
splitIndex <- createDataPartition(urb2$tc, p = 0.75, list = FALSE)
train_data <- urb2[splitIndex,]
test_data <- urb2[-splitIndex,]

model2 <- randomForest(tc ~ ., data = train_data, ntree = 100, mtry = 3)
print(model2)

```




Variable Importnace

```{r}

# Skapar graferna
res2            <- data.frame(varImp(model1)) # hämtar variable importance
res2$impProcent <- round(res2$Overall/sum(res2$Overall)*100) # räknar  om importance för varje till procent av hela
res2$var        <- rownames(res2) # kopierar radnamnet

res2 %>% ggplot(aes(x=impProcent, y=reorder(var, impProcent))) +
  geom_bar(stat="identity")



```

```{r}

# Skapar graferna
res2            <- data.frame(varImp(model2)) # hämtar variable importance
res2$impProcent <- round(res2$Overall/sum(res2$Overall)*100) # räknar  om importance för varje till procent av hela
res2$var        <- rownames(res2) # kopierar radnamnet

res2 %>% ggplot(aes(x=impProcent, y=reorder(var, impProcent))) +
  geom_bar(stat="identity")



```

#https://stackoverflow.com/questions/72234213/retain-all-columns-after-using-group-by-summarise-and-mutatue-dplyr-on-categori


LP

```{r}

library(lpSolve)

waste_data=df %>% group_by(region) %>% summarise(csor=mean(csor),
                                              cres=mean(cres), pop=mean(pop),msw=mean(msw))

sample_municipality <- waste_data %>% 
  slice(5) 

# Extract necessary values from the sample municipality
csor <- sample_municipality$csor
cres <- sample_municipality$cres
pop <- sample_municipality$pop
msw <- sample_municipality$msw

# Calculate costs
cost_sort <- csor/pop
cost_landfill <- cres/pop
cost_W2E <- 1.5 * cost_sort

# Define the objective function
f.obj <- c(cost_sort, cost_landfill, cost_W2E)

# Define the constraints
# The constraints are set up to ensure that the sum of the waste processed by all methods 
# equals the total waste, and that the waste processed by each method doesn't exceed their respective limits.
f.con <- matrix(c(1, 1, 1, 
                  0, 0, -1, 
                  0, -1, 0), 
                ncol=3, byrow=TRUE)

f.dir <- c("==", "<=", "<=")
f.rhs <- c(msw, -0.4 * msw, -0.5 * msw)

# Solve the LP problem
solution <- lp(direction="min", f.obj, f.con, f.dir, f.rhs, compute.sen=TRUE)

# Print the results https://stackoverflow.com/questions/72234213/retain-all-columns-after-using-group-by-summarise-and-mutatue-dplyr-on-categori

print(solution$solution)
print(solution$objval)


```





